---
title: "2000 census data"
author: "roberta barnett"
date: "10/19/2018"
output:
  pdf_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(GISTools)
#year 2000
#source https://s4.ad.brown.edu/Projects/Diversity/Researcher/LTBDDload/DataList.aspx
#codebook https://s4.ad.brown.edu/Projects/Diversity/Researcher/LTBDDload/Dfiles/codebooks.pdf
```

```{r}
# reading in data files 
## 2000 census data for the whole country
census2000 <- read.csv("~/GitHub/SpatialProject/data/2000/census2000.csv", sep=",")
```

```{r}
# cleaned Brooklyn data 
LTDB_Std_2000_fullcount <- read.csv("~/GitHub/SpatialProject/data/LTDB_Std_2000_fullcount/LTDB_Std_2000_fullcount.csv")

census2000 <- subset(LTDB_Std_2000_fullcount, LTDB_Std_2000_fullcount$state=="NY")
census2000 <- subset(LTDB_Std_2000_fullcount, LTDB_Std_2000_fullcount$county=="Kings County")
census2000. <- subset(census2000, census2000$state=="NY")

names(census2000.)[names(census2000.) == "POP00"] <- "population"
names(census2000.)[names(census2000.) == "NHWHT00"] <- "white"
names(census2000.)[names(census2000.) == "NHBLK00"] <- "black"
names(census2000.)[names(census2000.) == "NTV00"] <- "nativeamerican"
names(census2000.)[names(census2000.) == "ASIAN00"] <- "asian"
names(census2000.)[names(census2000.) == "HISP00"] <- "hispanic"
names(census2000.)[names(census2000.) == "HAW00"] <- "nativehawaiian"

write.csv(census2000., file="census2000.csv")

#shapefile of bk tracts
library(tigris)
Bk2000 <- tracts(state = 'NY', county = c('Kings'), year = 2000)

census2000.$ctlabel <- gsub("Census Tract ","",census2000.$tract)
```



```{r merging wb 2006 data with 2000 census tract data}
# reading 2000 census data
census2000_williamsburg <- read.csv("~/GitHub/SpatialProject/WB subproject/census2000_williamsburg.csv")

# creating ct labels for 2000
williamsburg_ctlabel2000 <- as.character(census2000_williamsburg$tract)
williamsburg_ctlabel2000 <- c(williamsburg_ctlabel2000, "515")
williamsburg_ctlabel2000 <- williamsburg_ctlabel2000[-3]

# subsetting...
Bk2000@data$ctlabel <- Bk2000@data$NAME00
nyc_census_wb_2000 <- Bk2000[Bk2000@data$ctlabel %in% williamsburg_ctlabel2000,]
```

```{r temporary data munging}
load(file="../../../../OneDrive/Documents/NYU/Fall 2018/Spatial Statistics/Datasets/Stop_Frisk/Stop_Frisk/sqf.RData")


stops_2006 <- stops[stops$year == 2006,]

## subsetting based on the bbox of willamsburg precincts
stops_2006_williamsburg <- stops_2006[stops_2006$lon > nyc_census_wb_2000@bbox[1,1] & stops_2006$lon < nyc_census_wb_2000@bbox[1,2] & stops_2006$lat < nyc_census_wb_2000@bbox[2,2] &
                                                  stops_2006$lat > nyc_census_wb_2000@bbox[2,1],]

## removing missing lat-lons
stops_2006_williamsburg <- stops_2006_williamsburg[!is.na(stops_2006_williamsburg$lat) & !is.na(stops_2006_williamsburg$lon),]

stops_2006_williamsburg <- as.data.frame(stops_2006_williamsburg)
coords <- stops_2006_williamsburg[,c("lon","lat")]
stops_2006_williamsburg_sp <- SpatialPointsDataFrame(coords=coords,data=stops_2006_williamsburg,
                                             proj4string=nyc_census_wb_2000@proj4string)

stops_census_merge <- gContains(nyc_census_wb_2000,stops_2006_williamsburg_sp,byid=TRUE)

tracts <- nyc_census_wb_2000$ctlabel
stops_2006_williamsburg$ctlabel <- NA
for(i in c(1:dim(stops_census_merge)[2])){
  stops_2006_williamsburg$ctlabel[stops_census_merge[,i]] <- tracts[i]
}

stops_subset <- stops_2006_williamsburg[!is.na(stops_2006_williamsburg$ctlabel),]
coords <- stops_subset[,c("lon","lat")]
stops_subset_sp <- SpatialPointsDataFrame(coords=coords,data=stops_subset,
                                             proj4string=nyc_census_wb_2000@proj4string)

# plotting
plot(nyc_census_wb_2000)
with(stops_subset, points(lon, lat, pch = ".", col = stops_subset$suspect.race))

```



```{r}
stops_subset$suspect.race <- stops_subset$suspect.race%>%fct_collapse(hispanic = c("black hispanic","white hispanic"))

stops_frisks_by_race <- stops_subset %>% group_by(ctlabel,suspect.race) %>% summarise(stops=n(),frisks=sum(frisked))

colnames(stops_frisks_by_race)[2] <- "group"

stops_frisks_total <- stops_subset %>% group_by(ctlabel) %>% summarise(stops=n(),frisks=sum(frisked))
stops_frisks_total$group <- "total"

stops_frisks_total <- stops_frisks_total[,c(1,4,2,3)]

stops_frisks_by_race <- rbind(as.data.frame(stops_frisks_by_race),as.data.frame(stops_frisks_total))

head(stops_frisks_by_race)

length(unique(stops_frisks_by_race$ctlabel))
```


```{r}
head(census2000.)
# names(census2000.NY)

census2000. <- census2000.[,c("population","white","black","nativeamerican","asian","hispanic","nativehawaiian","ctlabel")]

```



```{r}
stop_frisk_rates <- merge(stops_frisks_by_race,census2000.,by="ctlabel")
head(stop_frisk_rates)

stop_frisk_rates_total


length(unique(stop_frisk_rates$ctlabel))
```

```{r}
stop_frisk_rates <- stop_frisk_rates %>% mutate(stop_rate=case_when(
  group == "total" ~ stops*1000/population,
  group == "white" ~ stops*1000/white,
  group == "black" ~ stops*1000/black,
  group == "asian" ~ stops*1000/asian,
  group == "hispanic" ~ stops*1000/hispanic),frisk_rate = case_when(
  group == "total" ~ frisks*1000/population,
  group == "white" ~ frisks*1000/white,
  group == "black" ~ frisks*1000/black,
  group == "asian" ~ frisks*1000/asian,
  group == "hispanic" ~ frisks*1000/hispanic))
```



```{r}
stop_frisk_rates_total <- stop_frisk_rates[stop_frisk_rates$group == "total",]
choropleth(nyc_census_wb_2000,stop_frisk_rates_total[match(nyc_census_wb_2000$ctlabel,stop_frisk_rates_total$ctlabel),"frisk_rate"],main="Frisk Rate in 2006")
```
```{r}
wb06.nb <- poly2nb(nyc_census_wb_2000, queen=TRUE)
wb06.lw <- nb2listw(wb06.nb)
geary.mc(stop_frisk_rates_total$frisk_rate,wb06.lw,nsim=999) 
```
```{r}
wb06.nb <- poly2nb(nyc_census_wb_2000, queen=FALSE)
wb06.lw <- nb2listw(wb06.nb)
geary.mc(stop_frisk_rates_total$frisk_rate,wb06.lw,nsim=999) 
```
```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total$frisk_rate, wb06.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_wb_2000,moran.loc[,"Pr(z > 0)"],shades)
```

```{r}
moran.plot(stop_frisk_rates_total$frisk_rate, wb06.lw)
```


```{r}
stop_frisk_rates_total <- stop_frisk_rates[stop_frisk_rates$group == "total",]
choropleth(nyc_census_wb_2000,stop_frisk_rates_total[match(nyc_census_wb_2000$ctlabel,stop_frisk_rates_total$ctlabel),"stop_rate"],main="Stop Rate in 2006")
```

```{r}
wb06.nb <- poly2nb(nyc_census_wb_2000, queen=TRUE)
wb06.lw <- nb2listw(wb06.nb)
geary.mc(stop_frisk_rates_total$stop_rate,wb06.lw,nsim=999) 
```



```{r}
wb06.nb <- poly2nb(nyc_census_wb_2000, queen=FALSE)
wb06.lw <- nb2listw(wb06.nb)
geary.mc(stop_frisk_rates_total$stop_rate,wb06.lw,nsim=999) 
```
```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total$stop_rate, wb06.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_wb_2000,moran.loc[,"Pr(z > 0)"],shades)
```

```{r}
moran.plot(stop_frisk_rates_total$stop_rate, wb06.lw)
```

```{r loading files needed}
# shapefile
nyc_census_tracts <- readOGR("../data/2010 Census Tracts/geo_export_efe53346-3b34-4624-afa5-3012c025c024.shp")
nyc_census_tracts@data$ctlabel <- as.character(nyc_census_tracts@data$ctlabel)

# generating census tract labels
willamsburg_ctlabel <- censustrack2010.bk$ctlabel
willamsburg_ctlabel <- as.character(willamsburg_ctlabel)

# williamsburg shapefile
nyc_census_williamsburg <- nyc_census_tracts[nyc_census_tracts@data$ctlabel %in% willamsburg_ctlabel,]


# cleaning- getting rid of the accidental Bronx tract
nyc_census_williamsburg <- nyc_census_williamsburg[nyc_census_williamsburg$boro_name != "Bronx",]

nyc_census_williamsburg <- nyc_census_williamsburg[nyc_census_williamsburg$boro_name != "Queens",]

# wb subset for of stop frisk data (williamsburg_stops)
load(file.choose())

## subsettign only 2011
wb_stops_2011 <- williamsburg_stops[williamsburg_stops$year==2011,]

##subset for cts
wb_stops_2011 <- wb_stops_2011[]
wb_stops_2011 <- wb_stops_2011[wb_stops_2011$boro_ct201 %in% boro_ct_2011$V1,]

# 2010 census data (censustrack2010), will subset later
#load("~/GitHub/SpatialProject/data/cleaned_census_data.RData")
load(file.choose())
```

```{r some plots}
plot(nyc_census_williamsburg)
with(wb_stops_2011, points(lon, lat, pch = ".", col = wb_stops_2011$suspect.race))
legend("bottomleft", levels(wb_stops_2011$suspect.race), 
       fill = wb_stops_2011$suspect.race,
       bty = "n",
       cex = .6)
```



```{r generating choropleth for 2010}
stops_subset_2011 <- wb_stops_2011

stops_subset_2011$suspect.race <- stops_subset_2011$suspect.race%>%fct_collapse(hispanic = c("black hispanic","white hispanic"))

stops_frisks_by_race.2011 <- stops_subset_2011 %>% group_by(boro_ct201,suspect.race) %>% summarise(stops=n(),frisks=sum(frisked))

colnames(stops_frisks_by_race.2011)[2] <- "group"

stops_frisks_total.2011 <- stops_subset_2011 %>% group_by(boro_ct201) %>% summarise(stops=n(),frisks=sum(frisked))
stops_frisks_total.2011$group <- "total"

stops_frisks_total.2011 <- stops_frisks_total.2011[,c(1,4,2,3)]

stops_frisks_by_race.2011 <- rbind(as.data.frame(stops_frisks_by_race.2011),as.data.frame(stops_frisks_total.2011))

head(stops_frisks_by_race.2011)

length(unique(stops_frisks_by_race.2011$boro_ct201))
```

```{r}
censustrack2010.bk <- censustrack2010
censustrack2010.bk <- censustrack2010.bk[censustrack2010.bk$borough=="Brooklyn",]
censustrack2010.bk$censustract <- as.character(censustrack2010.bk$censustract)

# trimming and renaming to conform with 2000 census var names
censustrack2010.bk <- censustrack2010.bk[,-c(1:3)]
censustrack2010.bk$ctlabel <- as.character(as.numeric(censustrack2010.bk$censustract)/100)

# subsetting
williamsburg_ctlabel.2011 <- c(449, 561, 527, 509, 553, 551, 539, 547, 531, 285.01, 481, 495, 499, 503, 505, 517, 533,569, 591, 593, 529, 511, 537, 545, 501, 515, 523, 525, 453, 485, 489, 493, 535, 477, 497, 543, 589, 555, 389, 491, 507,519, 557, 513, 549)

censustrack2010.bk <- censustrack2010.bk[censustrack2010.bk$ctlabel %in% williamsburg_ctlabel.2011,]



# trimming and renaming again
censustrack2010.bk <- censustrack2010.bk[,-c(1)]
names(censustrack2010.bk)[1] <- "population"
names(censustrack2010.bk)[4] <- "nativeamerican"
censustrack2010.bk <- censustrack2010.bk[,c("population", "white", "black", 
                                            "nativeamerican", "asian", "hispanic",
                                            "nativehawaiian", "ctlabel")]
```


```{r}
#stop_frisk_rates.2011 <- merge(censustrack2010.bk,stops_frisks_total.2011,by.x="ctlabel", by.y="ctlabel")

stop_frisk_rates.2011 <- cbind(censustrack2010.bk, stops_frisks_total.2011)


head(stop_frisk_rates.2011)
```

```{r}
stop_frisk_rates.2011 <- stop_frisk_rates.2011 %>% mutate(stop_rate=case_when(
  group == "total" ~ stops*1000/population,
  group == "white" ~ stops*1000/white,
  group == "black" ~ stops*1000/black,
  group == "asian" ~ stops*1000/asian,
  group == "hispanic" ~ stops*1000/hispanic),frisk_rate = case_when(
  group == "total" ~ frisks*1000/population,
  group == "white" ~ frisks*1000/white,
  group == "black" ~ frisks*1000/black,
  group == "asian" ~ frisks*1000/asian,
  group == "hispanic" ~ frisks*1000/hispanic))
```

```{r}
stop_frisk_rates_total.2011 <- stop_frisk_rates.2011[stop_frisk_rates.2011$group == "total",]
stop_frisk_rates_total.2011[!is.na(stop_frisk_rates_total.2011$ctlabel),]

choropleth(nyc_census_williamsburg,stop_frisk_rates_total.2011[match(nyc_census_williamsburg$ctlabel,stop_frisk_rates_total.2011$ctlabel),"frisk_rate"],main="Frisk Rate in 2011")
```

```{r}
#geary's c
wb11.nb <- poly2nb(nyc_census_williamsburg, queen=TRUE)
wb11.lw <- nb2listw(wb11.nb)
geary.mc(stop_frisk_rates_total.2011$frisk_rate,wb11.lw,nsim=999) 
```
```{r}
#geary's c
wb11.nb <- poly2nb(nyc_census_williamsburg, queen=FALSE)
wb11.lw <- nb2listw(wb11.nb)
geary.mc(stop_frisk_rates_total.2011$frisk_rate,wb11.lw,nsim=999) 
```
```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total.2011$frisk_rate, wb11.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_williamsburg,moran.loc[,"Pr(z > 0)"],shades)
```
```{r}
moran.plot(stop_frisk_rates_total.2011$frisk_rate, wb11.lw)
```


```{r}
stop_frisk_rates_total.2011 <- stop_frisk_rates.2011[stop_frisk_rates.2011$group == "total",]
#stop_frisk_rates_total.2011[!is.na(stop_frisk_rates_total.2011$ctlabel),]

choropleth(nyc_census_williamsburg,stop_frisk_rates_total.2011[match(nyc_census_williamsburg$ctlabel,stop_frisk_rates_total.2011$ctlabel),"stop_rate"],main="Stop Rate in 2011")
```

```{r}
#geary's c
wb11.nb <- poly2nb(nyc_census_williamsburg, queen=TRUE)
wb11.lw <- nb2listw(wb11.nb)
geary.mc(stop_frisk_rates_total.2011$stop_rate,wb11.lw,nsim=999) 
```

```{r}
#geary's c
wb11.nb <- poly2nb(nyc_census_williamsburg, queen=FALSE)
wb11.lw <- nb2listw(wb11.nb)
geary.mc(stop_frisk_rates_total.2011$stop_rate,wb11.lw,nsim=999) 
```

```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total.2011$stop_rate, wb11.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_williamsburg,moran.loc[,"Pr(z > 0)"],shades)
```

```{r}
moran.plot(stop_frisk_rates_total.2011$stop_rate, wb11.lw)
```
#agg by race
#2006
stop_frisk_rates_total$blackrate <- 1000 * stop_frisk_rates_total$black/stop_frisk_rates_total$stops

stop_frisk_rates_total$whiterate <- 1000 * stop_frisk_rates_total$white/stop_frisk_rates_total$stops

stop_frisk_rates_total$hisprate <- 1000 * stop_frisk_rates_total$hispanic/stop_frisk_rates_total$stops

#2011

stop_frisk_rates_total.2011$blackrate <- 1000 * stop_frisk_rates_total.2011$black/stop_frisk_rates_total.2011$stops

stop_frisk_rates_total.2011$whiterate <- 1000 * stop_frisk_rates_total.2011$white/stop_frisk_rates_total.2011$stops

stop_frisk_rates_total.2011$hisprate <- 1000 * stop_frisk_rates_total.2011$hispanic/stop_frisk_rates_total.2011$stops


##blacks 


```{r}
choropleth(nyc_census_wb_2000,stop_frisk_rates_total[match(nyc_census_wb_2000$ctlabel,stop_frisk_rates_total$ctlabel),"blackrate"],main="Stop Rate in 2006 (Blacks)")
```
```{r}
#geary's c
geary.mc(stop_frisk_rates_total$black,wb06.lw,nsim=999) 
```

```{r}
choropleth(nyc_census_williamsburg,stop_frisk_rates_total.2011[match(nyc_census_williamsburg$ctlabel,stop_frisk_rates_total.2011$ctlabel),"blackrate"],main="Stop Rate in 2011 (Blacks)")
```

```{r}
#geary's c
wb11.nb <- poly2nb(nyc_census_williamsburg, queen=TRUE)
wb11.lw <- nb2listw(wb11.nb)
geary.mc(stop_frisk_rates_total.2011$blackrate,wb11.lw,nsim=999) 
```
##hispanics 

```{r}
choropleth(nyc_census_wb_2000,stop_frisk_rates_total[match(nyc_census_wb_2000$ctlabel,stop_frisk_rates_total$ctlabel),"hisprate"],main="Stop Rate in 2006 (Hispanics)")
```
```{r}
#geary's c
geary.mc(stop_frisk_rates_total$hispanic,wb06.lw,nsim=999) 
```



```{r}
choropleth(nyc_census_williamsburg,stop_frisk_rates_total.2011[match(nyc_census_williamsburg$ctlabel,stop_frisk_rates_total.2011$ctlabel),"hisprate"],main="Stop Rate in 2011 (Hispanics)")
```

```{r}
#geary's c
geary.mc(stop_frisk_rates_total.2011$hispanic,wb11.lw,nsim=999) 
```

##whites

```{r}
choropleth(nyc_census_wb_2000,stop_frisk_rates_total[match(nyc_census_wb_2000$ctlabel,stop_frisk_rates_total$ctlabel),"whiterate"],main="Stop Rate in 2006 (Whites)")
```
```{r}
#geary's c
geary.mc(stop_frisk_rates_total$white,wb06.lw,nsim=999) 
```

```{r}
choropleth(nyc_census_williamsburg,stop_frisk_rates_total.2011[match(nyc_census_williamsburg$ctlabel,stop_frisk_rates_total.2011$ctlabel),"white"],main="Stop Rate in 2011 (Whites)")
```

```{r}
#geary's c
geary.mc(stop_frisk_rates_total.2011$white,wb11.lw,nsim=999) 
```

```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total$blackrate, wb06.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_wb_2000,moran.loc[,"Pr(z > 0)"],shades)
```


```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total$whiterate, wb06.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_wb_2000,moran.loc[,"Pr(z > 0)"],shades)
```
```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total$hisprate, wb06.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_wb_2000,moran.loc[,"Pr(z > 0)"],shades)
```

```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total.2011$whiterate, wb11.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_williamsburg,moran.loc[,"Pr(z > 0)"],shades)
```
```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total.2011$hisprate, wb11.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_williamsburg,moran.loc[,"Pr(z > 0)"],shades)
```

```{r}
#localmoran
moran.loc <- localmoran(stop_frisk_rates_total.2011$blackrate, wb11.lw, p.adjust.method = "fdr")
shades <- shading(c(0.5),cols=c(2,8))
choropleth(nyc_census_williamsburg,moran.loc[,"Pr(z > 0)"],shades)
```

#mean stop and frisk rates

mean(stop_frisk_rates_total$stops)

mean(stop_frisk_rates_total$blackrate)

mean(stop_frisk_rates_total$whiterate) 

mean(stop_frisk_rates_total$hisprate) 

mean(stop_frisk_rates_total.2011$stops)

mean(stop_frisk_rates_total.2011$blackrate) 

mean(stop_frisk_rates_total.2011$whiterate)

mean(stop_frisk_rates_total.2011$hisprate)
